Estimating Learning Dynamics in the Routing Game

### 2015-10-13 paper/paper.tex
Added a conclusion

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\begin{definition}[Nash equilibrium]
A distribution $x^\star = (x_1^\star, \dots, x_K^\star)$ is a Nash equilibrium if it satisfies the following condition: for all other distributions $x = (x_1, \dots, x_K)$ and for all~$k$,
\[
\braket{\ell_k(x^\star)}{x_k - x_k^\star} \geq 0.
\]
\end{definition}
In words, $x^\star$ is a Nash equilibrium if for every player $k$, the expected cost under $x_k^\star$ is lower than the expected cost under any other distribution $x_k$. If we define the inner product $\braket{x}{\ell} = \sum_{k} \braket{x_k}{\ell_k}$, then this is equivalent to: $x^\star$ is an equilibrium if and only if $\braket{\ell(x^\star)}{x - x^\star} \geq 0$ for all feasible $x$. This variational inequality is, in fact, equivalent to the first-order optimality condition of the following potential function, usually referred to as the Rosenthal potential, in reference to~\cite{rosenthal1973class}:
\begin{proposition}[Rosenthal potential]
Consider a routing game and define the following function
\[
f(x) = \sum_{e \in E} \int_0^{\phi_e(x)} c_e(u)du.
\]
Then $f$ is convex its gradient is $\nabla f(x) = \ell(x)$.
\end{proposition}
This result can be found for example in~\cite{roughgarden2002bad}. Due to the fact that the loss function $\ell(\cdot)$ coincides with the gradient field $\nabla f(\cdot)$ of the Rosenthal potential, the Nash condition can be rewritten as
\[
\braket{\nabla f(x^\star)}{x - x^\star} \geq 0 \ \forall \text{ feasible }x,
\]
and since $f$ is convex, this is a necessary and sufficient condition for optimality of $x^\star$ (see e.g. Section 4.2.3 in~\cite{boyd2010convex}). Therefore the set of Nash equilibria is exactly the set of minimizers of the convex potential~$f$. This is important both for computation (computing a Nash equilibrium can be done by minimizing a convex function), and for modeling: one can model player dynamics as performing a distributed optimization of the potential function. More precisely, if we adopt the point of view presented in the introduction, wherein each player faces a sequential decision problem, and plays $x^{(t)}_k$ then observes $\ell_k(x^{(t)})$, then this corresponds to a first-order distributed optimization of the function~$f$, where each player is responsible for updating the variables $x_k^{(t)}$, and observes, at each iteration, the partial gradient $\ell_k(x^{(t)}) = \nabla_{x_k} f(x^{(t)})$. Using this connection to distributed optimization, a model of player dynamics was proposed in~\cite{krichene2015MD}. We review the model in the next Section.

@@

The convergence of this distributed learning model is discussed in~\cite{krichene2015MD}. In particular, the learning dynamics is guaranteed to converge under the following assumptions:
\begin{theorem}[Theorem~3 in~\cite{krichene2015MD}]
\label{thm:convergence}
Consider the routing game with mirror descent dynamics defined in Algorithm~\ref{alg:MD}, and suppose that for all $k$, $\eta_k^{(t)}$ is decreasing and converges to $0$. Then $f(x^{(t)}) - f^\star = \Ocal\parenth{\sum_k \frac{1}{t\eta_k^{(t)}} + \frac{\sum_{\tau = 1}^t \eta_k^{(\tau)}}{t} }$.
\end{theorem}

@@

%============================================================================================
\section{Conclusion}
\label{sec:conclusion}
Conclusion goes here.

%============================================================================================
%ACKNOWLEDGMENTS are optional
\section*{Acknowledgments}
Acknowledgement goes here.

%============================================================================================
%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\bibliographystyle{abbrv}
\bibliography{bib}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
%============================================================================================
\appendix
%Appendix A

Appendix goes here.

>>>>

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\begin{definition}[Nash equilibrium]
A distribution $x^\star = (x_1^\star, \dots, x_K^\star)$ is a Nash equilibrium if it satisfies the following condition: for all other feasible  distributions $x = (x_1, \dots, x_K)$ and for all~$k$,
\[
\braket{\ell_k(x^\star)}{x_k - x_k^\star} \geq 0.
\]
\end{definition}
In words, $x^\star$ is a Nash equilibrium if for every player $k$, the expected cost under $x_k^\star$ is lower than the expected cost under any other distribution $x_k$. If we define the inner product $\braket{x}{\ell} = \sum_{k} \braket{x_k}{\ell_k}$, then this is equivalent to: $x^\star$ is an equilibrium if and only if $\braket{\ell(x^\star)}{x - x^\star} \geq 0$ for all feasible $x$. This variational inequality is, in fact, equivalent to the first-order optimality condition of the following potential function, usually referred to as the Rosenthal potential, in reference to~\cite{rosenthal1973class}:
\begin{proposition}[Existence of a convex potential]
Consider a routing game and define the following function
\[
f(x) = \sum_{e \in E} \int_0^{\phi_e(x)} c_e(u)du.
\]
Then $f$ is convex its gradient is $\nabla f(x) = \ell(x)$.
\end{proposition}
This result can be found for example in~\cite{roughgarden2002bad}. Due to the fact that the loss function $\ell(\cdot)$ coincides with the gradient field $\nabla f(\cdot)$ of the Rosenthal potential, the Nash condition can be rewritten as
\[
\braket{\nabla f(x^\star)}{x - x^\star} \geq 0 \ \forall \text{ feasible }x,
\]
and since $f$ is convex, this is a necessary and sufficient condition for optimality of $x^\star$ (see e.g. Section 4.2.3 in~\cite{boyd2010convex}). Therefore the set of Nash equilibria is exactly the set of minimizers of the convex potential~$f$. This is important both for computation (computing a Nash equilibrium can be done by minimizing a convex function), and for modeling: one can model player dynamics as performing a distributed optimization of the potential function. More precisely, if we adopt the point of view presented in the introduction, wherein each player faces a sequential decision problem, and plays $x^{(t)}_k$ then observes $\ell_k(x^{(t)})$, then this corresponds to a first-order distributed optimization of the function~$f$, where each player is responsible for updating the variables $x_k^{(t)}$, and observes, at each iteration, the partial gradient $\ell_k(x^{(t)}) = \nabla_{x_k} f(x^{(t)})$. Using this connection to distributed optimization, a model of player dynamics was proposed in~\cite{krichene2015MD}. We review the model in the next Section.

@@

The convergence of this distributed learning model is discussed in~\cite{krichene2015MD}. In particular, the learning dynamics is guaranteed to converge under the following assumptions:
\begin{theorem}[Theorem~3 in~\cite{krichene2015MD}]
\label{thm:convergence}
Consider the routing game with mirror descent dynamics defined in Algorithm~\ref{alg:MD}, and suppose that for all $k$, $\eta_k^{(t)}$ is decreasing to~$0$. Then $f(x^{(t)}) - f^\star = \Ocal\parenth{\sum_k \frac{1}{t\eta_k^{(t)}} + \frac{\sum_{\tau = 1}^t \eta_k^{(\tau)}}{t} }$.
\end{theorem}

@@

%============================================================================================
\section{Conclusion}
\label{sec:conclusion}
We proposed a problem of model estimation in the routing game, to fit a distributed learning model to sequential observations of player decisions. The estimated model can then be used to predict the decisions at future iterations, or, more generally, as a plant model in an optimal control problem.

We considered in particular a model based on the mirror descent algorithm, parameterized by a DGF $\psi_k$ and a sequence of learning rates $(\eta^{(t)}_k)$, and gave an intuitive interpretation of how this model can describe player behavior. We showed that the problem of estimating one term of the learning rate sequence is convex in the case of the KL divergence (it remains open to prove this result for general Bregman divergences). To control the complexity of the model and to make the estimation consistent with the theoretical assumptions (decreasing learning rates), we proposed to parameterize the sequence with an initial term $\eta_k^{(0)}$ and a decay rate $\alpha_k \in (0, 1)$. When we tested these methods on data collected from our routing game interface, the parameterized sequence estimation outperformed the single term estimation on the prediction task. Our test results suggest that the mirror descent model can be a good descriptive model of player behavior, although in some rare cases, a player decision can be hard to model (e.g. when a player increase traffic assignment to previously bad routes).

This estimation problem can be extended in several ways: first, in our method, we fixed the DGF to be the negative entropy (regularized in order to avoid situations in which the estimation problem is ill-posed). One could also estimate the DGF itself, in addition to estimating the learning rates. One natural way to pose the estimation problem is to consider a finite collection of distance generating functions $\{\psi_i\}_{i \in \Ical}$, then to assume that each player $k$ uses a linear combination with weights $\theta_k$ $\psi = \sum_{i} \theta_{k, i} \psi_i$, then estimate the parameter vector $\theta_k$. One other natural approach is to pose the problem as a Bayesian estimation problem, by setting a prior on the model parameters (e.g. $\eta^{(0)}_k$ and $\alpha_k$), then define a likelihood model of observing a player decision given her DGF and learning rate.

%============================================================================================
%
% The following two commands are all you need in the
% initial runs of your .tex file to
% produce the bibliography for the citations in your paper.
\newpage
\bibliographystyle{abbrv}
\bibliography{bib}  % sigproc.bib is the name of the Bibliography in this case
% You must have a proper ".bib" file
%  and remember to run:
% latex bibtex latex latex
% to resolve all references
%
% ACM needs 'a single self-contained file'!
%
%APPENDICES are optional
%\balancecolumns
%============================================================================================
